\textbf{COMP0086 Summative Assignment}

### 2022-11-13 submission.tex
add constants.py to submission
edit submission and comment solutions

\item[(a)] Our sample space for images is $\{0, 1\}^D$, where each of our D dimensions can only take binary values, D being the number of pixels in the image.
The exponential family best suited on this sample space is the D-dimensional multivariate Bernoulli distribution because it shares the same sample space.
On the other hand, a D-dimensional multivariate Gaussian has the sample space $\mathbb{R}^D$, which does not match the sample space of our data.
To match our data sample space, we might have to define an additional mapping between our data and model spaces, adding unnecessary complexity.
Thus it would be inappropriate to model this dataset of images with a multivariate Gaussian.

\item[(b)] For $\{\textbf{x}^{(n)}\}_{n=1}^N$, a data set of N images, the joint likelihood (assuming images are independently and identically distributed) is the product of N D-dimensional multivariate Bernoulli distributions, one for each image:
$$P(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p}) = \prod_{n=1}^N P(\textbf{x}^{(n)} | \textbf{p})$$
Substituting the D-dimensional multivariate Bernoulli:
$$P(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p}) = \prod_{n=1}^{N}\prod_{d=1}^{D} p_d^{x_d^{(n)}} (1-p_d)^{1-x_d^{(n)}}$$
Taking the logarithm of this, we get the log likelihood:
$$\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p}) = \sum_{n=1}^{N}\sum_{d=1}^{D} [x_d^{(n)}\log(p_d) +  (1-x_d^{(n)})\log(1-p_d)]$$
Note that since the logarithm is a monotonically increasing function on $\mathbb{R}_+$, the maximisers and minimisers of the likelihood do not change.
Thus, to solve for the maximum likelihood estimate, $\hat p_d$, we can take the derivative of $\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})$ with respect to $p_d$, the $d^{th}$ element of $\textbf{p}$:
$$\frac{\partial\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})}{\partial p_d} = \sum_{n=1}^{N} (\frac{x_d^{(n)}}{p_d} -  \frac{1-x_d^{(n)}}{1-p_d})$$
$$\frac{\partial\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})}{\partial p_d} = \frac{\sum_{n=1}^{N} x_d^{(n)}}{p_d} -  \frac{\sum_{n=1}^{N} (1-x_d^{(n)})}{1-p_d}$$
and set the derivative to zero and solve for $\hat p_d$:
$$\frac{\sum_{n=1}^{N} x_d^{(n)}}{\hat p_d} -  \frac{\sum_{n=1}^{N} (1-x_d^{(n)})}{1-\hat p_d} = 0$$
$$ \sum_{n=1}^{N} x_d^{(n)} - \hat p_d\sum_{n=1}^{N} x_d^{(n)} - \hat p_d  \cdot N + \hat p_d \sum_{n=1}^{N}x_d^{(n)} = 0$$
$$  \hat p_d = \frac{1}{N}\sum_{n=1}^{N} x_d^{(n)}$$
Moreover, the second derivative with respect to $p_d$:
$$\frac{\partial\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})}{\partial p_d^2} = \frac{-\sum_{n=1}^{N} x_d^{(n)}}{p_d^2} +  \frac{-\sum_{n=1}^{N} (1-x_d^{(n)})}{(1-p_d)^2}$$
For a maximum, we need to show that the second derivative is negative.
Since $x_d^{(n)} \in \{0, 1\}$, in the worst case, of $N=1$, the single pixel $x_d^{(1)}$ must either be white ($\sum_{n=1}^{N} x_d^{(n)}> 0$) or black ($\sum_{n=1}^{N} 1-x_d^{(n)} > 0$) with the other being zero,  $\frac{\partial\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})}{\partial p_d^2} < 0$ will be guaranteed and $\hat p_d$ is a maximum as required for the maximum likelihood estimate.

@@

We know from (b):
$$\frac{\partial\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})}{\partial p_d} = \frac{\sum_{n=1}^{N} x_d^{(n)}}{p_d} -  \frac{\sum_{n=1}^{N} (1-x_d^{(n)})}{1-p_d}$$
For the second term $\frac{\partial\mathcal{L}(\textbf{p})}{\partial p_d}$, we start with $P(\textbf{p}))$, assuming each pixel to have an independent prior:
$$P(\textbf{p}) = \prod_{d=1}^D P(p_d)$$
Assuming a Beta prior on each $p_d$:
$$P(\textbf{p}) = \prod_{d=1}^D \frac{1}{B(\alpha, \beta)} p^{\alpha-1}_d (1-p_d)^{\beta-1}$$
Taking the logarithm:
$$\mathcal{L}(\textbf{p}) = \sum_{d=1}^{D} -\log (B(\alpha, \beta)) + (\alpha-1)\log p_d + (\beta-1)\log(1-p_d)$$
Taking the derivative with respect to $p_d$:
$$\frac{\partial\mathcal{L}(\textbf{p})}{\partial p_d} = \frac{(\alpha-1)}{p_d} -  \frac{(\beta-1)}{1-p_d}$$
Since we are only concerned with $p_d$, we are only left with a single element of the summation pertaining to $p_d$.

Combining, we have have an expression for $\frac{\partial\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N)}{\partial p_d}$:
$$\frac{\partial\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N)}{\partial p_d} = \frac{\sum_{n=1}^{N} x_d^{(n)}}{p_d} -  \frac{\sum_{n=1}^{N} (1-x_d^{(n)})}{1-p_d} + \frac{(\alpha-1)}{p_d} -  \frac{(\beta-1)}{1-p_d}$$
$$\frac{\partial\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N)}{\partial p_d} = \frac{(\alpha-1) + \sum_{n=1}^{N} x_d^{(n)}}{p_d} -  \frac{(\beta-1) + \sum_{n=1}^{N} (1-x_d^{(n)})}{1-p_d}$$
To find the maximum a posteriori (MAP) estimate $\hat{p_d}$ set $\frac{\partial\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N)}{\partial p_d} = 0$ and solve:
$$0 = \frac{(\alpha-1) + \sum_{n=1}^{N} x_d^{(n)}}{\hat{p_d}} -  \frac{(\beta-1) + \sum_{n=1}^{N} (1-x_d^{(n)})}{1-\hat{p_d}}$$
$$0 = (1-\hat{p_d})(\alpha-1) + (1-\hat{p_d})\bigg(\sum_{n=1}^{N} x_d^{(n)}\bigg) -  \hat{p_d}(\beta-1) - \hat{p_d}\bigg(\sum_{n=1}^{N} (1-x_d^{(n)})\bigg)$$
$$0 = (\alpha-\alpha \hat{p_d} + \hat{p_d} - 1) +\bigg(\sum_{n=1}^{N} x_d^{(n)} - \hat{p_d} \sum_{n=1}^{N} x_d^{(n)}\bigg) -  (\hat{p_d}\beta-\hat {p_d}) - \bigg(\hat{p_d}\cdot N - \hat{p_d}\sum_{n=1}^{N}x_d^{(n)}\bigg)$$
Cancelling the $\hat p_d\sum_{n=1}^{N} x_d^{(n)}$ terms:
$$0 = \alpha-\alpha \hat{p_d} + \hat{p_d} - 1 +\sum_{n=1}^{N} x_d^{(n)} -  \hat{p_d}\beta+\hat{p_d} - \hat{p_d}\cdot N$$
$$0 = \hat{p_d}(2-\alpha-\beta-N) + \alpha - 1 +\sum_{n=1}^{N} x_d^{(n)}$$
$$\hat{p_d} =  \frac{\alpha - 1 +\sum_{n=1}^{N} x_d^{(n)}}{(N+\alpha+\beta-2)}$$
To show that this is a maximum, the second derivative is:
$$\frac{\partial^2\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N)}{(\partial p_d)^2} = \frac{(1-\alpha) - \sum_{n=1}^{N} x_d^{(n)}}{(p_d)^2} +  \frac{(1-\beta) - \sum_{n=1}^{N} (1-x_d^{(n)})}{(1-p_d)^2}$$.
For a maximum, we need $\frac{\partial^2\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N)}{(\partial p_d)^2} < 0$ meaning that we need at least one of the strict inequalities $\alpha < 1- \sum_{n=1}^{N} x_d^{(n)}$ or $\beta < 1-\sum_{n=1}^{N} (1-x_d^{(n)})$ to be satisified, where the other can be $\leq$. The Beta distribution requires $\alpha > 0$ and $\beta > 0$ so this requirement will always be satisfied (in the worst case of a single image, either $x_d^{(1)}=1$ or $1-x_d^{(1)}=1$).

@@

Comparing the equations:
    $$\hat{\textbf{p}}^{MLE} = \frac{1}{N}\sum_{n=1}^{N} \textbf{x}^{(n)}$$
    and
    $$\hat{\textbf{p}}^{MAP}  =  \frac{\alpha - 1 +\sum_{n=1}^{N} \textbf{x}^n}{(N+\alpha+\beta-2)}$$
As the number of data points increases, $\hat{\textbf{p}}^{MAP}$ approaches $\frac{1}{N}\sum_{n=1}^{N} \textbf{x}^{(n)}$, $\hat{\textbf{p}}^{MLE}$.
This makes sense because as our data set gets bigger, we are less reliant on our prior.
However, if a specific pixel in all of the images of our data set are white or all black, the MLE for that pixel would either be 1 or 0.
This may not be representative of our intuitions about images, as there should be some non-zero probability of a pixel being black or white.
By introducing an appropriate prior we can ensure that the probability of that pixel will never be exactly zero or one.
In our case, with a Beta(3,3) prior on each pixel, our parameter values are biased to be closer to 0.5 and to never be at the extremities 0 and 1.
We can see this in Figure $\ref{fig:1e}$ where the range of our parameters is smaller than the range of Figure $\ref{fig:1d}$ and doesn't include zero.
Figure $\ref{fig:1e-comparison}$ visualises $\hat{\textbf{p}}^{MAP}-\hat{\textbf{p}}^{MLE}$ and we can see that for likelihoods greater than 0.5 in the MLE, the MAP has a lower value and for likelihoods less than 0.5, the MAP has a higher value, confirming our intuitions.

@@

Priors can also help ensure numerical stability during calculations.
The logarithm of zero is negative infinity, so having if the MLE is zero it can be problematic for log-likelihoods calculations whereas MAP can ensure non-zero probabilities.
Interestingly, when $\alpha=\beta=1$, $\hat{\textbf{p}}^{MLE} = \hat{\textbf{p}}^{MAP}$.
This is when the prior is a uniform distribution and so there is uniform bias on the location of $\textbf{p}$ and we recover the MLE.

On the other hand, a mis-specified prior can be problematic, as the estimated parameters might be skewed by the prior and not properly represent the underlying data generating process, this can result in parameter estimates that are worse than using the MLE if our data set is limited.
%In this example, we're using images of different digits as part of our data set. If a particular digit occurs more frequently in the data set, we will see that digit when visualising the MLE or MAP (the digit zero in this case), they simply provides the "average" image pixels. With a set of images with more equal representation of each digit, the parameter visualisation could be an image overlaying the different digits, which is not representative of the "most likely" image in the data. In this case where our data is likely clustered, using a multimodal likelihood is probably a better idea than the unimodal likelihood we've used.

@@

When all D components are generated from a Bernoulli distribution with $p_d = 0.5$, we have the likelihood function for model $M_1$:
$$P(\textbf{x}^{(n)|\textbf{p}^{(1)}}=[0.5, 0.5,...,0.5]^T, M_1) = \prod_{n=1}^{N}\prod_{d=1}^{D} (0.5)^{x_d^{(n)}} (0.5)^{1-x_d^{(n)}}$$
When all D components are generated from Bernoulli distributions with unknown, but identical, $p_d$, we have the likelihood function for model $M_2$:
$$P(\textbf{x}^{(n)}|\textbf{p}^{(2)}}=[p_d, p_d, ..., p_d]^T, M_2) = \prod_{n=1}^{N}\prod_{d'=1}^{D} p_d^{x_{d'}^{(n)}} (1-p_d)^{1-x_{d'}^{(n)}}$$
When each component is Bernoulli distributed with separate, unknown $p_d$, we have the likelihood function for model $M_3$:
$$P(\textbf{x}^{(n)}|\textbf{p}^{(3)}}=[p_1, p_2, ..., p_D]^T, M_3) = \prod_{n=1}^{N}\prod_{d=1}^{D} p_d^{x_d^{(n)}} (1-p_d)^{1-x_d^{(n)}}$$
For each model $M_i$, we can marginalise out $\textbf{p}^{(i)}}$ to get $P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)$:
$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i) = \int_0^1...\int_0^1 P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|p_d, M_i)P(p_d|M_i) dp_1...dp_D$$
where $d=1,...,D$ and $ \{\textbf{x}^{(n)}\}_{n=1}^{N}$ is our data set.

Given that the prior of any unknown probabilities is uniform, i.e. $P(p_d | M_i) = 1$. We can simplify:
$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i) = \int_0^1...\int_0^1 P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|p_d, M_i) dp_1...dp_D$$
For $M_1$, we have that all pixels have probability 0.5:
$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_1) = \int_0^1...\int_0^1 \prod_{n=1}^{N} \prod_{d=1}^{D} (0.5)^{x_d^{(n)}} (1-0.5)^{1-{x_d}^{(n)}} d\theta_1...d\theta_D$$
We can remove the integrals and knowing that either $x_d^{(n)}$ or $1-x_d^{(n)}$ will be 1 and the other zero, we can simplify $(0.5)^{x_d^{(n)}} (1-0.5)^{1-{x_d}^{(n)}}$ to $0.5$:
$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_1) = \prod_{n=1}^{N} \prod_{d=1}^D (0.5)$$
$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_1) = (0.5)^{N\cdot D}$$
For $M_2$, we have that all pixels share some probability $p_d$ so we only need to integrate over a single variable $p_d$:
$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \int_0^1 \prod_{n=1}^{N} \prod_{d'=1}^D p_d^{x_{d'}^{(n)}} (1-p_d)^{1-{x_{d'}}^{(n)}} dp_d$$
Changing the products to sums:
$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \int_0^1  p_d^{\sum_{n=1}^{N} \sum_{d'=1}^D x_{d'}^{(n)}} (1-p_d)^{\sum_{j=1}^{N} \sum_{d'=1}^D 1-{x_{d'}^{(n)}}} d p_d$$
Rewriting:
$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \int_0^1  (p_d)^{K} (1- p_{d'=1})^{N\cdot D-K} d p_d$$
where $K =\sum_{n=1}^{N} \sum_{d'=1}^D x_{d'}^{(n)}$.

This integral is the beta function:
$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \frac{K! (N\cdot D-k)!}{(N\cdot D+1)!}$$
For $M_3$, we need an integral for each $p_d$:
$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_3) = \int_0^1 ... \int_0^1 \prod_{n=1}^{N} \prod_{d=1}^D  p_d^{x_d^{(n)}} (1- p_d)^{1-{x_d}^{(n)}} d p_1 ... d p_D$$
We can separate the integrals to only contain the relevant $p_d$:
$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_3) = \prod_{d=1}^D \Bigg( \int_0^1 \prod_{n=1}^{N}   p_d^{x_d^{(n)}} (1- p_d)^{1-{x_d}^{(n)}} d p_d \Bigg)$$
Changing the products to sums:
$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_3) = \prod_{d=1}^D \Bigg( \int_0^1   p_d^{\sum_{n=1}^{N} x_d^{(n)}} (1- p_d)^{\sum_{n=1}^{N} 1-{x_d}^{(j)}} d p_d \Bigg)$$
In this case, we have the product of integrals where each evaluates to a beta function:
$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_3) = \prod_{d=1}^D \frac{K_d! (N-K_d)!}{(N+1)!}$$
where $K_d = \sum_{n=1}^{N} x_d^{(n}$.

>>>>

\item[(a)] Our sample space for images is $\{0, 1\}^D$, where each of our D dimensions can only take binary values (D being the number of pixels in the image).
The exponential family best suited on this sample space is the D-dimensional multivariate Bernoulli distribution because it shares the same sample space.
On the other hand, a D-dimensional multivariate Gaussian has the sample space $\mathbb{R}^D$, which does not match the sample space of our data.
%To match our data sample space, we might have to define an additional mapping between our data and model spaces, adding unnecessary complexity.
It is not immediately clear how the likelihood of an image of binary (discrete) values would be calculated under the continuous distribution of a multivariate Gaussian.
Thus it would be inappropriate to model this dataset of images with a multivariate Gaussian.

\item[(b)] For $\{\textbf{x}^{(n)}\}_{n=1}^N$, a data set of N images, the joint likelihood (assuming images are independently and identically distributed) is the product of N, D-dimensional multivariate Bernoulli distributions:
$$P(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p}) = \prod_{n=1}^N P(\textbf{x}^{(n)} | \textbf{p})$$
Substituting the D-dimensional multivariate Bernoulli:
$$P(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p}) = \prod_{n=1}^{N}\prod_{d=1}^{D} p_d^{x_d^{(n)}} (1-p_d)^{1-x_d^{(n)}}$$
Taking the logarithm, we get the log likelihood:
$$\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p}) = \sum_{n=1}^{N}\sum_{d=1}^{D} [x_d^{(n)}\log(p_d) +  (1-x_d^{(n)})\log(1-p_d)]$$
Note that since the logarithm is a monotonically increasing function on $\mathbb{R}_+$, the maximisers and minimisers of the likelihood do not change.
Thus, to solve for the maximum likelihood estimate, $\hat p_d$, we can take the derivative of $\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})$ with respect to $p_d$, the $d^{th}$ element of $\textbf{p}$:
$$\frac{\partial\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})}{\partial p_d} = \sum_{n=1}^{N} (\frac{x_d^{(n)}}{p_d} -  \frac{1-x_d^{(n)}}{1-p_d})$$
$$\frac{\partial\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})}{\partial p_d} = \frac{\sum_{n=1}^{N} x_d^{(n)}}{p_d} -  \frac{\sum_{n=1}^{N} (1-x_d^{(n)})}{1-p_d}$$
and set the derivative to zero to solve for $\hat p_d$:
$$\frac{\sum_{n=1}^{N} x_d^{(n)}}{\hat p_d} -  \frac{\sum_{n=1}^{N} (1-x_d^{(n)})}{1-\hat p_d} = 0$$
$$ \sum_{n=1}^{N} x_d^{(n)} - \hat p_d\sum_{n=1}^{N} x_d^{(n)} - \hat p_d  \cdot N + \hat p_d \sum_{n=1}^{N}x_d^{(n)} = 0$$
$$  \hat p_d = \frac{1}{N}\sum_{n=1}^{N} x_d^{(n)}$$
%Moreover, the second derivative with respect to $p_d$:
%$$\frac{\partial\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})}{\partial p_d^2} = \frac{-\sum_{n=1}^{N} x_d^{(n)}}{p_d^2} +  \frac{-\sum_{n=1}^{N} (1-x_d^{(n)})}{(1-p_d)^2}$$
%For a maximum, we need to show that the second derivative is negative.
%Since $x_d^{(n)} \in \{0, 1\}$, in the worst case, of $N=1$, the single pixel $x_d^{(1)}$ must either be white ($\sum_{n=1}^{N} x_d^{(n)}> 0$) or black ($\sum_{n=1}^{N} 1-x_d^{(n)} > 0$) with the other being zero, so  $\frac{\partial\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})}{\partial p_d^2} < 0$ will be guaranteed and $\hat p_d$ is a maximum as required for the maximum likelihood estimate.

@@

We know from (b):
$$\frac{\partial\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})}{\partial p_d} = \frac{\sum_{n=1}^{N} x_d^{(n)}}{p_d} -  \frac{\sum_{n=1}^{N} (1-x_d^{(n)})}{1-p_d}$$
For the second term $\frac{\partial\mathcal{L}(\textbf{p})}{\partial p_d}$, we start with $P(\textbf{p})$, assuming each pixel to have an independent prior:
$$P(\textbf{p}) = \prod_{d=1}^D P(p_d)$$
and assuming a Beta prior on each $p_d$:
$$P(\textbf{p}) = \prod_{d=1}^D \frac{1}{B(\alpha, \beta)} p^{\alpha-1}_d (1-p_d)^{\beta-1}$$
Taking the logarithm:
$$\mathcal{L}(\textbf{p}) = \sum_{d=1}^{D} -\log (B(\alpha, \beta)) + (\alpha-1)\log p_d + (\beta-1)\log(1-p_d)$$
Taking the derivative with respect to $p_d$:
$$\frac{\partial\mathcal{L}(\textbf{p})}{\partial p_d} = \frac{(\alpha-1)}{p_d} -  \frac{(\beta-1)}{1-p_d}$$
Since we are only concerned with $p_d$, we are only left with a single element of the summation pertaining to $p_d$.

Combining, we have have an expression for $\frac{\partial\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N)}{\partial p_d}$:
$$\frac{\partial\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N)}{\partial p_d} = \frac{\sum_{n=1}^{N} x_d^{(n)}}{p_d} -  \frac{\sum_{n=1}^{N} (1-x_d^{(n)})}{1-p_d} + \frac{(\alpha-1)}{p_d} -  \frac{(\beta-1)}{1-p_d}$$
$$\frac{\partial\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N)}{\partial p_d} = \frac{(\alpha-1) + \sum_{n=1}^{N} x_d^{(n)}}{p_d} -  \frac{(\beta-1) + \sum_{n=1}^{N} (1-x_d^{(n)})}{1-p_d}$$
To find the maximum a posteriori (MAP) estimate $\hat{p_d}$ set $\frac{\partial\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N)}{\partial p_d} = 0$ and solve:
$$0 = \frac{(\alpha-1) + \sum_{n=1}^{N} x_d^{(n)}}{\hat{p_d}} -  \frac{(\beta-1) + \sum_{n=1}^{N} (1-x_d^{(n)})}{1-\hat{p_d}}$$
$$0 = (1-\hat{p_d})(\alpha-1) + (1-\hat{p_d})\bigg(\sum_{n=1}^{N} x_d^{(n)}\bigg) -  \hat{p_d}(\beta-1) - \hat{p_d}\bigg(\sum_{n=1}^{N} (1-x_d^{(n)})\bigg)$$
$$0 = (\alpha-\alpha \hat{p_d} + \hat{p_d} - 1) +\bigg(\sum_{n=1}^{N} x_d^{(n)} - \hat{p_d} \sum_{n=1}^{N} x_d^{(n)}\bigg) -  (\hat{p_d}\beta-\hat {p_d}) - \bigg(\hat{p_d}\cdot N - \hat{p_d}\sum_{n=1}^{N}x_d^{(n)}\bigg)$$
Cancelling the $\hat p_d\sum_{n=1}^{N} x_d^{(n)}$ terms:
$$0 = \alpha-\alpha \hat{p_d} + \hat{p_d} - 1 +\sum_{n=1}^{N} x_d^{(n)} -  \hat{p_d}\beta+\hat{p_d} - \hat{p_d}\cdot N$$
$$0 = \hat{p_d}(2-\alpha-\beta-N) + \alpha - 1 +\sum_{n=1}^{N} x_d^{(n)}$$
$$\hat{p_d} =  \frac{\alpha - 1 +\sum_{n=1}^{N} x_d^{(n)}}{(N+\alpha+\beta-2)}$$
%To show that this is a maximum, the second derivative is:
%$$\frac{\partial^2\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N)}{(\partial p_d)^2} = \frac{(1-\alpha) - \sum_{n=1}^{N} x_d^{(n)}}{(p_d)^2} +  \frac{(1-\beta) - \sum_{n=1}^{N} (1-x_d^{(n)})}{(1-p_d)^2}$$.
%For a maximum, we need $\frac{\partial^2\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N)}{(\partial p_d)^2} < 0$ meaning that we need at least one of the strict inequalities $\alpha < 1- \sum_{n=1}^{N} x_d^{(n)}$ or $\beta < 1-\sum_{n=1}^{N} (1-x_d^{(n)})$ to be satisified, where the other can be $\leq$. The Beta distribution requires $\alpha > 0$ and $\beta > 0$ so this requirement will always be satisfied (in the worst case of a single image, either $x_d^{(1)}=1$ or $1-x_d^{(1)}=1$).

@@

Comparing the equations:
    $$\hat{\textbf{p}}^{MLE} = \frac{1}{N}\sum_{n=1}^{N} \textbf{x}^{(n)}$$
    and
    $$\hat{\textbf{p}}^{MAP}  =  \frac{\alpha - 1 +\sum_{n=1}^{N} \textbf{x}^n}{(N+\alpha+\beta-2)}$$
As the number of data points increases, $\hat{\textbf{p}}^{MAP}$ approaches $\frac{1}{N}\sum_{n=1}^{N} \textbf{x}^{(n)}$, the $\hat{\textbf{p}}^{MLE}$.
This makes sense because as our data set gets bigger, the effect of the prior diminishes.
However, if a specific pixel in all of the images of our data set are white or all black, the MLE for that pixel would either be 1 or 0.
This may not be representative of our intuitions about images, as there should be some non-zero probability of a pixel being black or white.
By introducing an appropriate prior we can ensure that the probability of that pixel will never be exactly zero or one.
In our case, with a Beta(3,3) prior on each pixel, our parameter values are biased to be closer to 0.5 and to never be at the extremities 0 and 1.
We can see this in Figure $\ref{fig:1e}$ where the range of our parameters is smaller than the range of Figure $\ref{fig:1d}$ and doesn't include zero.
Figure $\ref{fig:1e-comparison}$ visualises $\hat{\textbf{p}}^{MAP}-\hat{\textbf{p}}^{MLE}$ and we can see that for likelihoods greater than 0.5 in the MLE, the MAP has a lower value and for likelihoods less than 0.5, the MAP has a higher value, confirming our intuitions.

@@

Priors can also help ensure numerical stability during calculations.
The logarithm of zero is negative infinity, so having if the MLE is zero it can be problematic for log-likelihood calculations whereas MAP can ensure non-zero probabilities.
Interestingly, when $\alpha=\beta=1$, $\hat{\textbf{p}}^{MLE} = \hat{\textbf{p}}^{MAP}$.
This is when the prior is a uniform distribution and so there is uniform bias on the location of $\textbf{p}$ and we recover the MLE.

On the other hand, a mis-specified prior can be problematic, as the estimated parameters might be skewed by the prior and not properly represent the underlying data generating process, this can result in parameter estimates that are `worse' than using the MLE if our data set is limited in size.
%In this example, we're using images of different digits as part of our data set. If a particular digit occurs more frequently in the data set, we will see that digit when visualising the MLE or MAP (the digit zero in this case), they simply provides the "average" image pixels. With a set of images with more equal representation of each digit, the parameter visualisation could be an image overlaying the different digits, which is not representative of the "most likely" image in the data. In this case where our data is likely clustered, using a multimodal likelihood is probably a better idea than the unimodal likelihood we've used.

@@

\item[] When all D components are generated from a Bernoulli distribution with $p_d = 0.5$, we have the likelihood function for model $M_1$:
$$P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\textbf{p}^{(1)}}=[0.5, 0.5,...,0.5]^T, M_1) = \prod_{n=1}^{N}\prod_{d=1}^{D} (0.5)^{x_d^{(n)}} (0.5)^{1-x_d^{(n)}}$$
Knowing that either $x_d^{(n)}$ or $1-x_d^{(n)}$ will be 1 and the other zero, we can simplify $(0.5)^{x_d^{(n)}} (1-0.5)^{1-{x_d}^{(n)}}$ to $0.5$:
$$P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\textbf{p}^{(1)}}=[0.5, 0.5,...,0.5]^T, M_1) = \prod_{n=1}^{N} \prod_{d=1}^D (0.5)$$
$$P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\textbf{p}^{(1)}}=[0.5, 0.5,...,0.5]^T, M_1) = 0.5^{N\cdot D}$$
When all D components are generated from Bernoulli distributions with unknown, but identical, $p_d$, we have the likelihood function for model $M_2$:
$$P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\textbf{p}^{(2)}}=[p_d, p_d, ..., p_d]^T, M_2) = \prod_{n=1}^{N}\prod_{d'=1}^{D} p_d^{x_{d'}^{(n)}} (1-p_d)^{1-x_{d'}^{(n)}}$$
When each component is Bernoulli distributed with separate, unknown $p_d$, we have the likelihood function for model $M_3$:
$$P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\textbf{p}^{(3)}}=[p_1, p_2, ..., p_D]^T, M_3) = \prod_{n=1}^{N}\prod_{d=1}^{D} p_d^{x_d^{(n)}} (1-p_d)^{1-x_d^{(n)}}$$
For each model $M_i$, we can marginalise out $\textbf{p}^{(i)}}$ to get $P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)$:
$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i) = \int_0^1...\int_0^1 P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|\textbf{p}^{(i)}, M_i)P(\textbf{p}^{(i)}|M_i) dp_1...dp_D$$
Given that the prior of any unknown probabilities is uniform, i.e. $P(\textbf{p}^{(i)} | M_i) = 1$. We can simplify:
$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i) = \int_0^1...\int_0^1 P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\textbf{p}^{(i)}, M_i) dp_1...dp_D$$
For $M_1$:
$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_1) = \int_0^1...\int_0^1 0.5^{N\cdot D} d\theta_1...d\theta_D$$
We can remove the integrals:
$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_1) = (0.5)^{N\cdot D}$$
For $M_2$, we have that all pixels share some probability $p_d$ so we only need to integrate over a single variable $p_d$:
$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \int_0^1 \prod_{n=1}^{N} \prod_{d'=1}^D p_d^{x_{d'}^{(n)}} (1-p_d)^{1-{x_{d'}}^{(n)}} dp_d$$
Changing the products to sums:
$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \int_0^1  p_d^{\sum_{n=1}^{N} \sum_{d'=1}^D x_{d'}^{(n)}} (1-p_d)^{\sum_{j=1}^{N} \sum_{d'=1}^D 1-{x_{d'}^{(n)}}} d p_d$$
Rewriting:
$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \int_0^1  (p_d)^{K} (1- p_{d'=1})^{N\cdot D-K} d p_d$$
where $K =\sum_{n=1}^{N} \sum_{d'=1}^D x_{d'}^{(n)}$.
